Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-14 surfaces-write-up/surface-write-up.tex
lemma on generation of ker(psi) for P^m

\begin{lem} \label{P-m-generators}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$.  Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$.  Then $R_D$ is generated in degree at most $\sum_{i=1}^n \ell_i a_i$.
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align*}
	\Sigma = \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}.
\end{align*}
We note that $\Sigma$ has extremal rays given by the lattice points $e_i$ corresponding to
\[
	\epsilon_i = \prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}
\]
for each $i\in \{1, \ldots n\}$.  Therefore, applying proposition \ref{prop:cone-generation}, we see $R_D$ is generated in degree less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

\begin{lem}\label{P-m-relations-psi}
$\ker(\psi)$ is generated in degree at most \todo{insert number}.
\end{lem}
\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(z_i - \beta_j)\prod_{j=1}^n {z_j}^{c_{i}}
\end{equation}

@@

To find such a $\beta$, we simply write $\epsilon_j$ as a 
polynomial in $\epsilon_1, \ldots \epsilon _3$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{i =
1}^n \epsilon_i^{r_i}$ to a canonical form. \todo{add reference to
earlier description of canonical form.}
\end{proof}

@@

\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(z_i - \beta_j)\prod_{j=1}^n {z_j}^{c_{i}}
\end{equation}

>>>>

\begin{lem} \label{P-m-generators}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$.  Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$.  Then $R_D$ is generated in degree at most $\sum_{i=1}^n \ell_i a_i$.
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}.
\end{align}
We note that $\Sigma$ has extremal rays given by the lattice points $e_i$ corresponding to
\begin{equation}\label{eqn:epilon-def-P-m}
	\epsilon_i = \prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}
\end{equation}
for each $i\in \{1, \ldots n\}$.  Therefore, applying proposition \ref{prop:cone-generation}, we see $R_D$ is generated in degree less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

\begin{lem}\label{P-m-relations-psi}
$\ker(\psi)$ is generated in degree at most 
\[
	\delta_i + \sum_{i=1}^n \ell_i a_i.
\]
\end{lem}
\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(z_i - \beta_j)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

@@

To find such a $\beta$, we simply write $\epsilon_j$ as a 
polynomial in $\epsilon_1, \ldots \epsilon _3$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{i =
1}^n \epsilon_i^{r_i}$ to a canonical form. \todo{add reference to
earlier description of canonical form.}

Now we bound the degree of generation of these relations for a fixed $z_i$.  We associate $(z_i - \beta_j)\prod_{j=1}^n z_j^{c_j}$ that has degree $d$ with the lattice point $(d, c_1, \ldots, c_n)$.  Let $\Sigma_i \subseteq \mathbb{R}^{n+1}$ be the set of points of the form $(d, r_1, \ldots, r_n)$ satisfying $r_j \ge -d \alpha_j$ for all $j$ and $\sum_{j=1}^n r_j a_j = -a_i$.  Then we see $\Sigma_i$ is a translation of $\Sigma$ (see Equation \ref{eqn:Sigma-def}) by $\delta_i = (\frac{a_i}{\deg(D)}, -\frac{\alpha_1 a_1}{\deg(D)}, \ldots, - \frac{\alpha_n a_n}{\deg(D)})$.

Therefore, we can write any element of $\Sigma'$ as
\[
	\delta_i + \sum_{j=1}^n r_j e_j
\]
for $r_j \in \mathbb{R}$ for each $j$.

Whenever some $r_j \ge 1$, we can generate the generation at hand by one of lower degree by factoring out the section $\epsilon_j$ corresponding to the lattice point $e_j$ (see Equation \ref{eqn:epilon-def-P-m}).  Therefore, we can bound the degree of generation of $\ker(\psi)$ by
\[
	\delta_i + \sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(z_i - \beta_j)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}
